\chap{Acknowledgment}

This dissertation would not have been possible without the guidance, help, and support of the following people.

First and foremost, I'd like to sincerely thank my Ph.D. advisor Dr. Soudeh Ghorbani for her insights and support throughout this journey. Under her guidance, I learned to be an independent researcher, and to effectively shape and present my research. I am forever grateful for this opportunity.

I'd like to thank Jean Tourrilhes who patiently mentored me during my internship at Hewlett Packard Labs. I learned a ton and more from him!

I'd like to thank my Master's degree advisor Dr. Mohsen Sharifi who taught me how to think about the philosophy of designing systems.

I'd like to thank all my co-authors. None of this has been possible without their help. I thank Dr. Puneet Sharma at Hewlett Packard Labs. There is nothing that can't happen with his unlimited support.
I sincerely thank Dr. Showan Asyabi and Dr. Alireza Sanaee. I started my research journey with them and have been ever asking for their insights. I thank my dear friend Sepehr Abdous for all his help and being there in all deadline crunches!
I thank Dr. Timothy Zhu and Raymond Matson. I learned a lot from them.
I thank the current and previous members of the Foundational Networked Systems Lab and the Network and Distributed Systems Lab at HPE for their feedback and support.

I'd like to recognize my Graduate Board and Thesis committee for their insights and comments on my research. I thank Dr. Yair Amir for the deep conversations that shaped my thinking when I started my Ph.D at Johns Hopkins. I thank Dr. John Ousterhout for the opportunity to help improve Homa implementation through rigorous measurement and experimentation.
I thank Dr. Alex Marder for the opportunity to help with the Computer Networks course along with all the advice on entering the job market. I thank Dr. Gianni Antichi for his warm and helpful comments on my research. I thank Dr. Ryan Huang for teaching me invaluable insights on the history and evolution of system software design. I thank Dr. Tamas Budavari for his comments and insights on earlier versions of my research.

Finally, I'd like to thank my family. My mom, dad, and my partner for their love and support. I am forever indebted to them.